HUME: Human UCCA-Based Evaluation of Machine Translation

### 2016-06-02 paper/main.tex
editting paper

\documentclass[11pt,letterpaper]{article}
%\usepackage[round]{natbib}
\usepackage{emnlp2016}
\usepackage{amsmath}
\usepackage{paralist}
\usepackage{subfig}
\usepackage{times}
\usepackage{latexsym}
\usepackage{graphicx}
\usepackage{url}
\usepackage{pgfplotstable}
%\usepackage{hyperref}
\usepackage{color}
\usepackage{lipsum,adjustbox}

@@

\newcommand{\com}[1]{}
\newcommand{\secref}[1]{Section~\ref{#1}}
\newcommand{\figref}[1]{Figure~\ref{#1}}
\newcommand{\tabref}[1]{Table~\ref{#1}}
\newcommand{\XXX}[1]{{\color{red}XXX #1}} % a general todo or comment
%\newcommand{\XXX}[1]{}
\newcommand{\oa}[1]{\footnote{\color{red}OA: #1}}
%\newcommand{\oa}[1]{}
\newcommand{\bh}[1]{\footnote{\color{blue}BH: #1}}
%\newcommand{\bh}[1]{}

@@

% better maths typography
\def\setsize#1{\lvert #1 \rvert}
\def\func#1{\text{\it #1}}  % much better kerning within function names
\def\HUME{\func{HUME}}
\def\Correct{\func{Correct}}
\def\Partial{\func{Partial}}
\def\Units{\func{Units}}

@@

Human evaluation of machine translation normally uses sentence-level measures such as relative ranking
or adequacy/fluency scales. However these provide no insight into the MT errors,
 and do not scale well with sentence 
length. We argue for a semantics-based evaluation, which aims to capture how much of the meaning
is retained in the MT output, and can support the construction and tuning of semantics-based MT. 
We present a novel human semantic evaluation measure for MT, Human
UCCA-based MT Evaluation (HUME), building on the Universal Conceptual Cognitive Annotation (UCCA) semantic representation scheme.
Making progress over previous work, the proposed measure takes into account
a wider range of semantic phenomena and does not rely on semantic annotation
of the MT output itself, whose interpretation is often unclear.
We experiment with four language pairs, demonstrating HUME's broad applicability,
as well as its reliability, reflected in its inter-annotator agreement rates and 
good correlation with human adequacy scores.

@@

Human judgement is the most important criterion for
estimating the quality of an MT system.
Nevertheless, common measures for human MT evaluation, such as adequacy and fluency judgements
or the relative ranking of possible translations, are problematic in two ways.
First, as the quality of translation is determined by multiple factors, it is difficult
to assign a single number that quantifies the quality of the entire sentence. This
is indeed reflected in the diminishing inter-annotator agreement (IAA) rates of human ranking measures
%as a function of
with
the sentence length \parcite{Bojar:2011}.
Second, a sentence-level quality score does not indicate what parts of the sentence
are badly translated, and so cannot inform developers in repairing these errors.

These problems are partially addressed by measures that decompose over parts of the evaluated
translation. For automatic measures,
these are often words or n-grams, for manual measures some structural
information is taken into account \parcite{machacek:bojar:segranks:2015}.
% the following is specific to automatic measures:
%quantifying the overlap of its sub-parts relative a reference translation.
A promising line of research decomposes metrics
%More recent work on MT evaluation proposed measures
%that decompose
over semantically defined units,
quantifying the similarity of the output and the reference in terms of
their verb argument structure; the most notable of these measures is
HMEANT \parcite{lo2011structured}.
% SAVING SPACE:
% \XXX{Do we want to cut the following shorter?}
% However, 
% their focus on verb argument structures misses many semantic phenomena
% (such as inter-clause linkage and nominal argument structures), providing
% only
% a limited perspective on the semantic similarity between the translation and reference.
% Also, assigning semantic structure to an MT output, especially a poor one,
% is difficult and arguably ill-defined.

@@

HUME operates by aggregating human assessments of the translation quality of individual
semantic units in the source sentence. We 
%HUME only requires semantically annotating the source sentence,
are thus avoiding the semantic annotation of machine-generated text. 
This allows the re-use of the source semantic annotation for measuring the quality
of different translations of the same source sentence.
%By basing the measure on a direct comparison
%of the source and translation we avoid the difficulties created by using
We also avoid creating and relying on possibly suboptimal
reference translations, albeit at the cost of requiring the
evaluator to be proficient in both the source and target languages.
%An elaborate comparison of HUME and HMEANT is given in \secref{sec:hmeant_comp}.

After a brief review (\secref{sec:background}), we describe HUME in detail
(\secref{sec:hume}).
Our experiments with four language pairs: English to Czech, German, Polish and Romanian (\secref{sec:experiments}) document the
efficiency (time of annotation) and inter-annotator agreement scores for HUME. 
We also report on the correlation scores of HUME with direct assesment of human 
adequacy ratings. 
Finally, the differences with HMEANT are discussed 
(\secref{sec:hmeant_comp}).

@@

\paragraph{Machine Translation Evaluation.}
%MT evaluation has been traditionally pursued through human-based
%evaluation that quantifies the adequacy of the translation in preserving the meaning 
%of the source sentence, and its fluency as a sentence in the target language
%\parcite{lopez2008statistical}.
Human evaluation is generally done by ranking the outputs of multiple systems
e.g., in the WMT tasks \parcite{bojar2015findings}, or by giving
adequacy/fluency scores to each translation, a procedure recently improved
by \perscite{graham2015improving}.
However, while providing the gold standard for MT evaluation, human evaluation
is not a scalable solution.
%and is infeasible where a wide variety of systems or parameter settings are to be evaluated.
% has motivated the construction of cheaper methods that correlate with human
%judgement.

@@

UCCA's current inventory of distinctions focuses on argument structures
(adjectival, nominal, verbal and others) and relations between them.
The most basic notion is the Scene, which describes a movement, an
action or a state which persists in time. Each Scene contains one main
relation and zero or more participants. For example, the sentence ``After graduation, Tom moved to NYC''
contains two Scenes, whose main relations are ``graduation'' and ``moved''.
The participant ``Tom'' is a part of both Scenes, while ``NYC'' only of the
latter (\figref{fig:ucca_example_v2}). Further categories account for
relations between Scenes and the internal structures of complex participants and relations.

% UCCA Motivation
UCCA is a natural candidate for defining a semantic
MT evaluation measure for a number of reasons. 
First, it defines an intuitive set of categories
that can be reliably annotated by non-experts, after as little as two hours
of training \parcite{marinotti2014}.
Second, UCCA is a cross-linguistically applicable scheme, seeking to
represent what is shared between languages
by building on linguistic typological theory, and specifically on ``Basic Linguistic Theory''
\parcite{Dixon:10a,Dixon:10b,Dixon:12}.
The cross-linguistic applicability of UCCA has been tested in annotations of
English, French, German and Czech.
Third, the scheme has been shown to be semantically stable
in translation: UCCA annotations of translated text usually contain the same set of relationships
\parcite{sulem2015conceptual}. This indicates that the scheme reflects
a common layer of representation that in a correct translation
is mostly shared between the translation and the source.

We note that verbal argument structures are less stable in this sense, as a verbal
argument structure may be translated or paraphrased to a nominal or adjectival
argument structure, and is thus less suitable as a basis for a semantic measure.
E.g., ``after he graduated, Tom moved to NYC'' contains two verbs,
but is similar in meaning to the one-verbed
``after graduation, Tom moved to NYC''.

@@

The Abstract Meaning Representation (AMR) scheme \parcite{banarescu2013abstract}
shares much of UCCA's motivation for defining a more complete semantic annotation.
However, using AMR is not optimal for defining a decomposition of a sentence into semantic
units as it does not ground its semantic symbols in the text,
thus does not provide clear decomposition of the sentence into sub-units.
Also, AMR is more fine-grained than UCCA and consequently harder to annotate.
Other approaches pursued a grounded representation of semantics,
and have generally represented semantic structures as bi-lexical dependencies.
Examples include the tectogrammatical layer \parcite{sgallhp:1986} used in the Prague dependency
treebanks, e.g. the Czech-English one \parcite{hajic2012announcing}, as well
as dependencies derived from Minimal Recursion Semantics representations \parcite{oepen2006discriminant}.
Nevertheless, despite their appeal as deep and intuitive means for semantic representation,
these approaches require linguistic expertise for their annotation, due to the intricate and
fine-grained distinctions they provide, and so we argue they are less suitable as a basis
for an interpretable MT evaluation.

@@

This section summarises the manual annotation procedure used
to compute the HUME measure. 
We denote the source sentence as $s$ and the translation as $t$. 
The procedure involves two manual steps: (1) UCCA-annotating $s$, 
(2) human judgements as to the translation quality of each semantic unit of $s$ relative to $t$,
where units are defined according to the UCCA annotation.
Note that UCCA annotation is performed once for every source sentence,
irrespective of the number of its translations we wish
to evaluate\footnote{Instead of relying on the source,
the evaluation could be based on the UCCA structure for the \emph{reference}
translation.
More details in the discussion in \secref{src-vs-ref}.
% This would allow monolingual annotators to take part in the
% evaluation. On the downside, the translation would have to be aligned to the
% reference and more importantly, we would risk inaccurate judgements due to
% the naturally occurring differences between the source and its reference
% translations.
}.

\paragraph{UCCA Annotation.}
We begin by creating UCCA annotations for the source sentence, following the
UCCA guidelines\footnote{All UCCA-related resources can be found
  here: \url{http://www.cs.huji.ac.il/~oabend/ucca.html}}.
Let $L$ be the set of possible UCCA labels.
Given a sentence $s = (w_1, \ldots,w_n)$, its UCCA annotation is a DAG $G=(V,E,\ell)$,
where $\ell: V \rightarrow L$, and where $W = \{w_1,\ldots,w_n\}$ correspond to the leaves of $G$.
For every node $v \in V$, we define its {\it yield} $yld(v)$ as $v$'s descendants in $W$.
The semantic units for a sentence $s$ according to $G$
is the set of yields of nodes in $V$. 
The translation is not UCCA-annotated.

@@

\paragraph{Translation Evaluation.}
The basic strategy for annotating HUME is to traverse the semantic units
of the source sentence, which correspond to the arguments and relations expressed
in the text, and to mark the extent to which they have been correctly translated.
HUME aggregates the judgements of the users into a composite score, 
which reflects the overall extent to which the semantic content of $s$ is preserved in $t$.

@@

Atomic units can be labelled as ``Green'' (correct), ``Orange'' (partially correct)
and ``Red'' (incorrect). 
A Green label means that the meaning of the word or phrase has been largely preserved.
An Orange label means that the essential meaning of the unit has been preserved,
but some part of the translation is wrong.
This could often be due to the translated word having the wrong tense,
or the wrong morphology, in a way that impacts little on the understandability of the sentence.
A Red label means that the essential meaning of the unit has not been captured.

Structural units have sub-units (children in the UCCA graph), which may themselves
either be atomic or structural.
% Ondrej thinks the following can be simply removed:
%The labels assigned to such nodes reflect the extent to which
%the relation between the sub-units is preserved in $t$.
%\XXX{(BH: don't follow this) For instance, a Scene unit represents the relation between the
%Scene elements, such as the predicate (main relation), participants and adverbials.}
Structural units are labeled as ``Adequate'' or ``Bad'', meaning
that the relation between the sub-units went wrong.
We use a three label system with atomic units, as
opposed to a two label one in structural units,
as atomic units often have slight errors, in which cases we do want to provide partial credit.

There are many ways that the relationship between the children might go
wrong in translation:

\begin{compactitem}
\item The components of the translation are ordered differently from the source,
  such that the relation between them changes. For instance, if
  the sentence ``man bites dog'' is translated into a sentence meaning ``dog bites man'',
  the semantic unit corresponding to the Scene will be marked as Bad, while
  the individual atomic units ``dog'', ``bites'' and ``man'' will be marked as Green.
\item A superfluous word or phrase has been inserted into
  the translation. For instance, if ``dog bites man'' is translated into ``dog viciously bites man''.
\item One of the sub-units is missing from the translation.
  For instance, if ``dog bites man'' is translated into ``dog bites''.
% XXX let's leave the reviewers to notice this:
%  \bh{Note that in this case we also 
%  get an error in the atomic unit, assuming that ``man'' is a semantic unit in the source, but for the other
%  two cases there is no atomic unit error.}
\end{compactitem}

@@

In non-configurational languages, the relationship between the
governing and dependent units is often expressed using morphological
properties of the dependants instead of their ordering
(this holds especially for the verb and its modifiers,
less so for components of noun phrases). Errors in morphology
should be expressed on the atomic units, so HUME can behave differently
on configurational vs. non-configurational languages.

@@

\begin{figure}
    \begin{center}
    \includegraphics[width=0.35\textwidth]{ucca-tree-mteval-v2.png}
    \end{center}
  \caption{\label{fig:hume_tree_v2}
     HUME annotation of an UCCA tree with a word aligned example translation shown below. 
Atomic units are labelled using traffic lights and structural nodes are marked A or B.}
\end{figure}
% Ondrej removes the following, he thinks that errors are not supposed to be
% propagated in HMEANT either:
%In HMEANT, it is not clear whether errors should be propegated up the SRL hierarchy.

In Figure~\ref{fig:hume_tree_v2} we can see an example of a possible HUME
annotation, with the translation shown in English for ease of comprehension.
When evaluating ``to NYC'' the annotator looks at the translation and sees the
word ``stateside''. This word captures the whole phrase and so we mark this
non-leaf node with an atomic label. Here we choose ``Orange'' since
it approximately captures the meaning
in this context. The ability to mark non-leaves with atomic labels allows
the annotator to account for translations which only correspond at the phrase
level. Another feature highlighted in this example is that by separating structural
and atomic nodes, we are able to define where an error occurs, and localise
the error to its point of origin. The linker ``After'' is translated incorrectly as ``by''
which changes the meaning of the entire sentence. This error is captured at
the atomic level, and it is labelled Red. The sentence still contains two scenes and
a linker and therefore we mark the root node as structurally correct, Adequate.

\begin{figure*}[t]
    \begin{center}
    \includegraphics[width=1\textwidth]{hume_interface2.jpg}
    \caption{The HUME annotation tool, based on the UCCA tool. The faded orange box at the top
      contains the translation, and directly below it is the source sentence. The tree of semantic units
      is shown below the source. Alignments between the source and translation appear in italics and unaligned intervening words appear in red.}
    \label{fig:interface}
    \end{center}
\end{figure*}

@@

We proceed to detailing how judgements on the semantic units
of the source are aggregated into a composite score. We start by taking a parsimonious approach
%While future extensive empirical evaluation with HUME and its application to MT
%system tuning may suggest more complex definition,
%as a starting point we take a parsimonious approach,
and compute an accuracy score.

Let $\Correct(s,t)$ be the number of atomic units in $s$ marked as Green relative to $t$ plus
the number of structural units in $s$ marked as Adequate relative to $t$.
Let $\Partial(s,t)$ be the number of atomic units in $s$ marked as Orange relative
to $t$. Let $\setsize{\Units(s)}$ be the total number of units marked with any of the five labels.
Then HUME's composite score is:

\vspace{-.5cm}
\begin{equation}
  \small
  \HUME(s,t) = \frac{\Correct(s,t) + 0.5\cdot \Partial(s,t)}{\setsize{\Units(s)}}
  \nonumber
\end{equation}

@@

The interface presents, for each unit, the translation segment aligned with it.
This allows the user, especially in long sentences, to focus her attention on the parts
most likely to be relevant for her judgement. As the alignments are automatically derived,
and therefore noisy, the annotator is instructed to treat the aligned text is a cue, but to ignore
the alignment if it is misleading, and instead 
make a judgement according to the full translation.

Concretely, let $s = (w_1,\ldots,w_n)$ be a source sentence, $t = (w'_1,\ldots,w'_n)$ a translation,
and $A \subset 2^{s \times t}$ a many-to-many word alignment between them.
If $u$ is a semantic unit in $s$, whose yield is $yld(u)$, we define the aligned text in
$t$ to be $a(u)=\bigcup_{(x_s,x_t) \in A \wedge x_s \cap yld(u) \neq \emptyset} x_t \subset t$.
Where $a(u)$ is discontinuous in the text, words between the left
and right boundaries of $a(u)$ which are not contained in it (intervening words)
are presented in a smaller red font.
Intervening words are likely to change the meaning of the translation
of $u$, and thus should be paid attention to when considering whether the translation
is correct or not.

For example, 
%consider the sentence $s=$``We're driving back on Friday'' and the German
%translation t=``Am Freitag fahren wir zur\"uck'' (lit. ``on Friday driving we back'').
%``driving back'' will be aligned to ``fahren ... zur\"uck'',
%and ``wir'' will be marked as intervening word. Another example
in \figref{fig:interface}, ``ongoing pregnancy'' is translated to
``Schwangerschaft ... laufenden'' (lit. ``pregnancy ... ongoing''). This alone
seems acceptable but the interleaving words in red notify the annotator to check
the whole translation. The meaning of the expression is not preserved in the
translation
and the annotator should mark this structural node as ``B'', bad. 
%The interleaving 
%German words shown in red should not occur there, and the annotator should take this
%into account and mark this structural node as ``B'', bad.

@@

In order to validate the HUME metric, we ran an annotation experiment with one source language (English),
and four target languages (Czech, German, Polish and Romanian), using text from the public health domain.
The domain is particularly suitable for experimenting on MT evaluation, as semantically
accurate translation is paramount here, underscoring the importance of reliable
semantic evaluation measures.

@@

The experiment seeks to address two questions:
\begin{inparaenum}
\item Can HUME be annotated quickly and reliably, and does the reliability still hold for longer sentences?
\item Can HUME capture translation issues that other annotation schemes (such as
adequacy judgements and HMEANT)
cannot?
%\item What can HUME annotation tell us about the strengths and weaknesses of MT?
%\bh{Maybe we should remove the third one, we
%probably don't address it in this paper}
\end{inparaenum}

\subsection{Data Sets and Translation Systems}

@@

The source sentences are all in English, and their UCCA annotation was performed by four
computational linguists and one linguist.

For the annotation of the MT output, we recruited two annotators for each of German, Romanian
and Polish and one main annotator for Czech.
%In order to allow us to calculate
For IAA also for
Czech, several further annotators worked on a small number of 
sentences each. We treat these further annotators as one annotator, giving us two annotators
for each language pair.
The annotators were all native  speakers of the respective target languages and fluent in English.

>>>>

\documentclass[11pt,letterpaper]{article}
%\usepackage[round]{natbib}
\usepackage{emnlp2016}
\usepackage{amsmath}
\usepackage{paralist}
\usepackage{subfig}
\usepackage{times}
\usepackage{latexsym}
\usepackage{graphicx}
\usepackage{url}
\usepackage{pgfplotstable}
%\usepackage{hyperref}
\usepackage{color}
\usepackage{lipsum,adjustbox}
\usepackage[font={small}]{caption}

@@

\newcommand{\com}[1]{}
\newcommand{\secref}[1]{\S\ref{#1}}
\newcommand{\figref}[1]{Figure~\ref{#1}}
\newcommand{\tabref}[1]{Table~\ref{#1}}
\newcommand{\XXX}[1]{{\color{red}XXX #1}} % a general todo or comment
%\newcommand{\XXX}[1]{}
%\newcommand{\oa}[1]{\footnote{\color{red}OA: #1}}
\newcommand{\oa}[1]{}
%\newcommand{\bh}[1]{\footnote{\color{blue}BH: #1}}
\newcommand{\bh}[1]{}

@@

% better maths typography
\def\setsize#1{\lvert #1 \rvert}
\def\func#1{\text{\it #1}}  % much better kerning within function names
\def\HUME{\func{HUME}}
\def\Adequate{\func{Adequate}}
\def\Green{\func{Green}}
\def\Orange{\func{Orange}}
\def\Units{\func{Units}}

@@

Human evaluation of machine translation normally uses sentence-level measures such as relative ranking
or adequacy/fluency scales. However these provide no insight into the MT errors,
and do not scale well with sentence length.
We argue for a semantics-based evaluation, which captures what meaning components
are retained in the MT output, providing a more fine-grained analysis of
translation quality, and enables the construction and tuning of semantics-based MT. 
We present a novel human semantic evaluation measure for MT, Human
UCCA-based MT Evaluation (HUME), building on the Universal Conceptual Cognitive Annotation (UCCA)
semantic representation scheme.
Making progress over previous work, the proposed measure takes into account
a wider range of semantic phenomena and does not rely on semantic annotation
of the MT output itself, whose interpretation is often unclear.
We experiment with four language pairs, demonstrating HUME's broad applicability,
as well as its reliability, reflected in its inter-annotator agreement rates and 
good correlation with human adequacy scores.

@@

Human judgement is the cornerstone for estimating the quality of an MT system.
Nevertheless, common measures for human MT evaluation, such as adequacy and fluency judgements
or the relative ranking of possible translations, are problematic in two ways.
First, as the quality of translation is multi-faceted, it is difficult
to quantify the quality of the entire sentence in a single number. This
is indeed reflected in the diminishing inter-annotator agreement (IAA) rates of human ranking measures
with the sentence length \cite{Bojar:2011}.
Second, a sentence-level quality score does not indicate what parts of the sentence
are badly translated, and so cannot inform developers in repairing these errors.

These problems are partially addressed by measures that decompose over parts of the evaluated
translation. For automatic measures,
these are often words or n-grams, for manual measures some structural
information is taken into account \parcite{machacek:bojar:segranks:2015}.
% the following is specific to automatic measures:
%quantifying the overlap of its sub-parts relative a reference translation.
A promising line of research decomposes metrics
%More recent work on MT evaluation proposed measures
%that decompose
over semantically defined units,
quantifying the similarity of the output and the reference in terms of
their verb argument structure; the most notable of these measures is HMEANT
\parcite{lo2011structured}.

% SAVING SPACE:
% \XXX{Do we want to cut the following shorter?}
% However, 
% their focus on verb argument structures misses many semantic phenomena
% (such as inter-clause linkage and nominal argument structures), providing
% only
% a limited perspective on the semantic similarity between the translation and reference.
% Also, assigning semantic structure to an MT output, especially a poor one,
% is difficult and arguably ill-defined.

@@

HUME operates by aggregating human assessments of the translation quality of individual
semantic units in the source sentence. We 
%HUME only requires semantically annotating the source sentence,
are thus avoiding the semantic annotation of machine-generated text,
which is often garbled or semantically unclear.
This also allows the re-use of the source semantic annotation for
measuring the quality of different translations of the same source sentence,
and 
%By basing the measure on a direct comparison
%of the source and translation we avoid the difficulties created by using
avoids relying on possibly suboptimal reference translations,
albeit at the cost of requiring the
evaluator to be proficient in both the source and target languages.
%An elaborate comparison of HUME and HMEANT is given in \secref{sec:hmeant_comp}.

After a brief review (\secref{sec:background}), we describe HUME in detail
(\secref{sec:hume}).
Our experiments with four language pairs: English to Czech, German, Polish and Romanian (\secref{sec:experiments})
document HUME's inter-annotator agreement and efficiency (time of annotation). 
We further empirically compare HUME with direct assesment of human adequacy ratings,
and conclude by discussing the differences with HMEANT (\secref{sec:hmeant_comp}).

@@

\paragraph{MT Evaluation.}
%MT evaluation has been traditionally pursued through human-based
%evaluation that quantifies the adequacy of the translation in preserving the meaning 
%of the source sentence, and its fluency as a sentence in the target language
%\parcite{lopez2008statistical}.
Human evaluation is generally done by ranking the outputs of multiple systems
e.g., in the WMT tasks \parcite{bojar2015findings}, or by giving
adequacy/fluency scores to each translation, a procedure recently improved
by \perscite{graham2015improving}.
However, while providing the gold standard for MT evaluation, human evaluation
is not a scalable solution.
%and is infeasible where a wide variety of systems or parameter settings are to be evaluated.
% has motivated the construction of cheaper methods that correlate with human
%judgement.

@@

UCCA's current inventory of distinctions focuses on argument structures
(adjectival, nominal, verbal and others) and relations between them.
The most basic notion is the Scene, which describes a movement, an
action or a state which persists in time. Each Scene contains one main
relation and zero or more participants. For example, the sentence ``After graduation, Tom moved to NYC''
contains two Scenes, whose main relations are ``graduation'' and ``moved''.
The participant ``Tom'' is a part of both Scenes, while ``NYC'' only of the
latter (\figref{fig:ucca_example_v2}). Further categories account for
inter-scene relations and the sub-structures of participants and relations.

% UCCA Motivation
The use of UCCA for semantic MT evaluation measure is motivated by several reasons. 
First, UCCA's set of categories can
be reliably annotated by non-experts after as little as two hours
of training \parcite{marinotti2014}.
Second, UCCA is cross-linguistically applicable, seeking to
represent what is shared between languages by building on
linguistic typological theory \parcite{Dixon:10a,Dixon:10b,Dixon:12}.
Its cross-linguistic applicability has so far been tested in annotations of
English, French, German and Czech.
Third, the scheme has been shown to be stable across translations:
UCCA annotations of translated text usually contain the same set of relations
\parcite{sulem2015conceptual}, indicating that the scheme reflects
a common layer of representation that in a correct translation
is mostly shared between the translation and the source\oa{verb argument structure note?}.

%Verbal argument structures are less stable in this sense, as they
%may be translated or paraphrased to a nominal or adjectival
%argument structure, and is thus less suitable as a basis for a semantic measure.
%E.g., ``after he graduated, Tom moved to NYC'' contains two verbs,
%but is similar in meaning to the one-verbed
%``after graduation, Tom moved to NYC''.

@@

The Abstract Meaning Representation (AMR) \parcite{banarescu2013abstract} project
shares UCCA's motivation for defining a more complete semantic annotation.
However, using AMR is not optimal for defining a decomposition of a sentence into semantic
units as it does not ground its semantic symbols in the text,
and thus does not provide clear decomposition of the sentence into sub-units.
Also, AMR is more fine-grained than UCCA and consequently harder to annotate.
Other approaches represent semantic structures as bi-lexical dependencies
\parcite{sgallhp:1986,hajic2012announcing,oepen2006discriminant},
which are indeed grounded in the text, but are less suitable for MT evaluation
as they require linguistic expertise for their annotation.

%, and so we argue they are less suitable 
%for an interpretable MT evaluation.

%Examples include the tectogrammatical layer  used in the Prague dependency
%treebanks, e.g. the Czech-English one \parcite{}, as well
%as dependencies derived from Minimal Recursion Semantics representations \parcite{}.
%Nevertheless, despite their appeal as deep and intuitive means for semantic representation,

@@

This section summarises the manual annotation procedure used
to compute the HUME measure. 
We denote the source sentence as $s$ and the translation as $t$. 
The procedure involves two manual steps: (1) UCCA-annotating $s$, 
(2) human judgements as to the translation quality of each semantic unit of $s$ relative to $t$,
where units are defined according to the UCCA annotation.
UCCA annotation is performed once for every source sentence,
irrespective of the number of its translations we wish
to evaluate\footnote{Instead of relying on the source,
the evaluation could be based on the UCCA structure for the \emph{reference}
translation.
More details in the discussion in \secref{src-vs-ref}.
% This would allow monolingual annotators to take part in the
% evaluation. On the downside, the translation would have to be aligned to the
% reference and more importantly, we would risk inaccurate judgements due to
% the naturally occurring differences between the source and its reference
% translations.
}.

\paragraph{UCCA Annotation.}
We begin by creating UCCA annotations for the source sentence, following the
UCCA guidelines\footnote{All UCCA-related resources can be found
  here: \url{http://www.cs.huji.ac.il/~oabend/ucca.html}}.
A UCCA annotation for a sentence $s$ is a labeled DAG $G$, whose leaves
are the words of $s$.
For every node in $G$, we define its {\it yield} to be its leaf descendants.
The semantic units for $s$ according to $G$ are the yields of nodes in $G$.

@@

\paragraph{Translation Evaluation.}
HUME annotation is done by traversing the semantic units
of the source sentence, which correspond to the arguments and relations expressed
in the text, and marking the extent to which they have been correctly translated.
HUME aggregates the judgements of the users into a composite score, 
which reflects the overall extent to which the semantic content of $s$ is preserved in $t$.

@@

Atomic units can be labelled as ``Green'' (correct), ``Orange'' (partially correct)
and ``Red'' (incorrect). 
Green means that the meaning of the word or phrase has been largely preserved.
Orange means that the essential meaning of the unit has been preserved,
but some part of the translation is wrong.
This is often be due to the translated word having the the wrong inflection,
in a way that impacts little on the understandability of the sentence.
Red means that the essential meaning of the unit has not been captured.

Structural units have sub-units (children in the UCCA graph), which may themselves
either be atomic or structural.
% Ondrej thinks the following can be simply removed:
%The labels assigned to such nodes reflect the extent to which
%the relation between the sub-units is preserved in $t$.
%\XXX{(BH: don't follow this) For instance, a Scene unit represents the relation between the
%Scene elements, such as the predicate (main relation), participants and adverbials.}
Structural units are labeled as ``Adequate'' or ``Bad'', meaning
that the relation between the sub-units went wrong\footnote{
  Three labels are used with atomic units, as opposed to two labels with structural units,
as atomic units are more susceptible to slight errors.}.
There are three major cases where the relationship between the children might go
wrong in translation.
First, the components of the translation may be ordered differently from the source,
such that the relation between them changes. For instance, if
the sentence ``man bites dog'' is translated into a sentence meaning ``dog bites man'',
the semantic unit corresponding to the Scene will be marked as Bad, while
the individual atomic units ``dog'', ``bites'' and ``man'' will be marked as Green.
Second, A superfluous word or phrase may be inserted into
the translation. For instance, if ``dog bites man'' is translated into ``dog viciously bites man''.
Third, one of the sub-units may be missing from the translation.
For instance, if ``dog bites man'' is translated into ``dog bites''.

@@

\com{
In non-configurational languages, the relationship between the
governing and dependent units is often expressed using morphological
properties of the dependants instead of their ordering
(this holds especially for the verb and its modifiers,
less so for components of noun phrases). Errors in morphology
should be expressed on the atomic units, so HUME can behave differently
on configurational vs. non-configurational languages.
}

@@

\begin{figure}
    \begin{center}
    \includegraphics[width=0.35\textwidth]{ucca-tree-mteval-v2.png}
    \end{center}
  \caption{\label{fig:hume_tree_v2}
     HUME annotation of an UCCA tree with a word aligned example translation shown below. 
Atomic units are labelled using traffic lights and structural units are marked A or B.}
\end{figure}
% Ondrej removes the following, he thinks that errors are not supposed to be
% propagated in HMEANT either:
%In HMEANT, it is not clear whether errors should be propegated up the SRL hierarchy.

Figure~\ref{fig:hume_tree_v2} presents an example of a HUME
annotation, where the translation is in English for ease of comprehension.
When evaluating ``to NYC'' the annotator looks at the translation and sees the
word ``stateside''. This word captures the whole phrase and so we mark this
non-leaf node with an atomic label. Here we choose Orange since
it approximately captures the meaning in this context.
The ability to mark non-leaves with atomic labels allows
the annotator to account for translations which only correspond at the phrase
level. Another feature highlighted in this example is that by separating structural
and atomic units, we are able to define where an error occurs, and localise
the error to its point of origin. The linker ``After'' is translated incorrectly as ``by''
which changes the meaning of the entire sentence. This error is captured at
the atomic level, and it is labelled Red. The sentence still contains two scenes and
a linker and therefore we mark the root node as structurally correct, Adequate.

\begin{figure*}[t]
    \begin{center}
    \includegraphics[width=1\textwidth]{hume_interface2.jpg}
    \caption{The HUME annotation tool. The top orange box
      contains the translation. The source sentence is directly below it, followed by the tree of the source
      semantic units. Alignments between the source and translation are in italics and
      unaligned intervening words are in red (see text).}
    \label{fig:interface}
    \end{center}
\end{figure*}

@@

We proceed to detailing how judgements on the semantic units
of the source are aggregated into a composite score. 
%While future extensive empirical evaluation with HUME and its application to MT
%system tuning may suggest more complex definition,
%as a starting point we take a parsimonious approach,
We start by taking a very simple approach and compute an accuracy score.
Let $\Green(s,t)$, $\Adequate(s,t)$ and $\Orange(s,t)$ be the number of Green, Adequate and Orange
units respectively. Let $\Units(s)$ be the number of units marked with any of the labels.
Then HUME's composite score is:

\vspace{-.5cm}
{\scriptsize
\begin{equation}
  \HUME(s,t) = \frac{\Green(s,t) + \Adequate(s,t) + 0.5\cdot \Orange(s,t)}{\Units(s)}
  \nonumber
\end{equation}
}
\vspace{-.5cm}

@@

The interface presents, for each unit, the translation segment aligned with it.
This allows the user, especially in long sentences, to focus her attention on the parts
most likely to be relevant for her judgement. As the alignments are automatically derived,
and therefore noisy, the annotator is instructed to treat the aligned text is a cue, but to ignore
the alignment if it is misleading, and instead 
make a judgement according to the full translation.
Concretely, let $s$ be a source sentence, $t$ a translation,
and $A \subset 2^{s \times t}$ a many-to-many word alignment.
If $u$ is a semantic unit in $s$, whose yield is $yld(u)$, we define the aligned text in
$t$ to be $\bigcup_{(x_s,x_t) \in A \wedge x_s \cap yld(u) \neq \emptyset} x_t$.

Where the aligned text is discontinuous in $t$, words between the left
and right boundaries which are not contained in it (intervening words)
are presented in a smaller red font. 
Intervening words are likely to change the meaning of the translation
of $u$, and thus should be paid attention to when considering whether the translation
is correct or not.

For example, 
%consider the sentence $s=$``We're driving back on Friday'' and the German
%translation t=``Am Freitag fahren wir zur\"uck'' (lit. ``on Friday driving we back'').
%``driving back'' will be aligned to ``fahren ... zur\"uck'',
%and ``wir'' will be marked as intervening word. Another example
in \figref{fig:interface}, ``ongoing pregnancy'' is translated to
``Schwangerschaft ... laufenden'' (lit. ``pregnancy ... ongoing''). This alone
seems acceptable but the interleaving words in red notify the annotator to check
the whole translation, in which the meaning of the expression is not preserved.
The annotator should thus mark this structural node as Bad.

%The interleaving 
%German words shown in red should not occur there, and the annotator should take this 
%into account and mark this structural node as ``B'', bad.

@@

In order to validate the HUME metric, we ran an annotation experiment with one source language (English),
and four target languages (Czech, German, Polish and Romanian), using text from the public health domain.
Semantically accurate translation is paramount in this domain, which makes
it particularly suitable for experimenting on MT evaluation. HUME is both evalauted in terms of its
consistency (inter-annotator agreement), efficiency (time of annotation) and validity (through a comparison
with crowd-sourced adequacy judgements).

@@

%The experiment seeks to address two questions:
%\begin{inparaenum}
%\item Can HUME be annotated quickly and reliably, and does the reliability still hold for longer sentences?
%\item Can HUME capture translation issues that other annotation schemes (such as adequacy judgements
%  and HMEANT) cannot?
%\end{inparaenum}

%\item What can HUME annotation tell us about the strengths and weaknesses of MT?
%\bh{Maybe we should remove the third one, we
%probably don't address it in this paper}

\subsection{Datasets and Translation Systems}

@@

The source sentences are all in English, and their UCCA annotation was performed by four
computational linguists and one linguist.
For the annotation of the MT output, we recruited two annotators for each of German, Romanian
and Polish and one main annotator for Czech. For Czech IAA,
several further annotators worked on a small number of 
sentences each. We treat these further annotators as one annotator, resulting in two annotators
for each language pair.
The annotators were all native speakers of the respective target languages and fluent in English.

\tabref{tab:annot}
shows the total number of sentences and units annotated by each annotator. Note
that not all units in all sentences were annotated, often
due to
%this was because
the annotator not finishing a sentence or accidentally missing a node.
%, but in many cases it was also because it was an implicit 
%UCCA node, which was not shown to the annotators since there is no corresponding source segment.
\begin{table}
\begin{center}
{\small
\begin{tabular}{ll|cccc}
& & cs & de & pl & ro \\
\hline
Sentences &  Annot. 1 & 324   & 339  & 351  & 230  \\
 & Annot. 2 & 205 & 104  & 340  & 337 \\
\hline
Units & Annot. 1 & 8794  & 9253 & 9557  & 6152 \\
 &Annot. 2 & 5553 & 2906  & 9303  & 9228  \\
\end{tabular}
\caption{Number of HUME-annotated sentences and units.}
\label{tab:annot}
}
\end{center}
\end{table}
